---
title: "quanteda.dictionaries"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/quanteda.dictionaries)](https://cran.r-project.org/package=quanteda.dictionaries)
[![Travis-CI Build Status](https://travis-ci.org/kbenoit/quanteda.dictionaries.svg?branch=master)](https://travis-ci.org/kbenoit/quanteda.dictionaries)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/kbenoit/quanteda.dictionaries?branch=master&svg=true)](https://ci.appveyor.com/project/kbenoit/quanteda.dictionaries) 
![Coverage Status](https://img.shields.io/codecov/c/github/kbenoit/quanteda.dictionaries/master.svg)](https://codecov.io/github/kbenoit/quanteda.dictionaries?branch=master)

An R package consisting of dictionaries for text analysis and associated utilities.  Designed to be used with [**quanteda**](http://quanteda.io) but can be used more generally with any text analytic package (e.g. **tidytext**, **tm**, etc.).

## Installing

```{r eval = FALSE}
# the devtools package needs to be installed for this to work
devtools::install_github("quanteda/quanteda") 
```


## Sources

* [AFINN](http://www2.imm.dtu.dk/pubdb/views/publication_details.php?id=6010)

* [NRC Word-Emotion Association Lexicon, also called EmoLex](http://saifmohammad.com/WebPages/lexicons.html)

* [Affective Norms for English Words (ANEW)](http://csea.phhp.ufl.edu/media/anewmessage.html)

* [vaderSentiment](https://github.com/cjhutto/vaderSentiment)

* [Bing and Liu](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html)

* [Other dictionaries at the Provalis page](https://provalisresearch.com/products/content-analysis-software/wordstat-dictionary/)

* [Lexicoder](http://www.lexicoder.com)

* [Moral Foundations Dictionary](http://www.moralfoundations.org/othermaterials)

* [Regressive Imagery Dictionary](http://www.kovcomp.co.uk/wordstat/RID.html)

* [Loughran and McDonald Financial Sentiment](http://www3.nd.edu/~mcdonald/Word_Lists.html) See also [here](http://sraf.nd.edu/textual-analysis/resources/).

* [Harvard IV-4 dictionaries](http://www.wjh.harvard.edu/~inquirer/homecat.htm)


## Formats

* quanteda

## Languages

Some dictionaries are available in multiple languages - the package should have a way to retrieve that.  Maybe a `summary()` method for the dictionaries that indicates the languages available?

## Utility functions

* `as.yaml()`  
* `as.list()`  
* `c.dictionary()`
* something to convert to the **tidytext** dictionary format

## Copyright information

Probably need to list the (c) from each dictionary here.

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
