---
title: "quanteda.dictionaries"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, results="hide", message=FALSE}
library("badger")
```

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/quanteda.dictionaries)](https://cran.r-project.org/package=quanteda.dictionaries)
`r badge_devel("kbenoit/quanteda.dictionaries", "royalblue")`
[![Travis-CI Build Status](https://travis-ci.org/kbenoit/quanteda.dictionaries.svg?branch=master)](https://travis-ci.org/kbenoit/quanteda.dictionaries)
[![AppVeyor Build status](https://ci.appveyor.com/api/projects/status/3uvg00bo4p7mr98s/branch/master?svg=true)](https://ci.appveyor.com/project/kbenoit/quanteda-dictionaries/branch/master)
[![Coverage Status](https://img.shields.io/codecov/c/github/kbenoit/quanteda.dictionaries/master.svg)](https://codecov.io/github/kbenoit/quanteda.dictionaries?branch=master)

An R package consisting of dictionaries for text analysis and associated utilities. Designed to be used with [**quanteda**](http://quanteda.io) but can be used more generally with any text analytic package (e.g. **tidytext**, **tm**, etc.).

## Installing

```{r eval = FALSE}
# the devtools package needs to be installed for this to work
devtools::install_github("kbenoit/quanteda.dictionaries") 
```

## Demonstration

With the `liwcalike()` function from the **quanteda.dictionaries** package, you can easily analyze text corpora using exising or custom dictionaries. Here we show how to apply the Lexicoder Sentiment Dictionary (Young and Soroka 2012) to a corpus consting of 2000 movie reviews (from the [**quanteda.textmodels**](https://github.com/quanteda/quanteda.textmodels) package).

```{r, warning=FALSE, message=FALSE}
library("quanteda.dictionaries")

# requires GitHub version of quanteda.textmodels
output_lsd <- liwcalike(quanteda.textmodels::data_corpus_moviereviews,
                        dictionary = data_dictionary_NRC)

head(output_lsd)
```

## Code of Conduct

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.
